hatchR: A toolset to predict hatch and emergence phenology in wild fishes

### 2024-12-11 inst/manuscript/paper.Rmd
clean up some code chunks to make things shorter
add fig.caps

rerender Rmd to PDF

```{r echo = FALSE, out.height = '50%', fig.align = 'center'}
knitr::include_graphics("flowchart.png")
```

@@

```{r}
plot_check_temp(data = year_sim,
                dates = date,
                temperature = temp,
                temp_min = 0, # temp_min and max lines are 
                              # user customizable
                temp_max = 25)
```

@@

```{r, eval=FALSE}
# it is useful to plot again to check your summarized data
plot_check_temp(data = year_sim_summ,
                dates = date,
                temperature = daily_temp,
                temp_min = 0, # temp_min and max lines are 
                              # user customizable
                temp_max = 15)
```

@@

fit_model for three non-salmonid specieslib

@@

Below, we demonstrate how the `fit_model()` function may be used to create custom parameterizations for species beyond the Salmonids in the `model_table` included in the package. We include parameterizations from three warm-water species to demonstrate the `fit_model()` utility for fishes beyond the scope of the original effective value approach. We include parameterizations for commonly cultured sportfishes including Smallmouth Bass (*Micropterus dolomieu*), Channel Catfish (*Ictalurus punctatus*) from @small2001, and Lake Sturgeon (*Acipenser fulvescens*) from @smith2005.

@@

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE}
### predict_phenology

@@

sockeye_emerge_mod <- model_select(
  author = "Beacham and Murray 1990", 
  species = "sockeye", 
  model = 2, 
  development_type = "emerge"
  )

@@

We can also do the same with emergence:

```{r, warning=FALSE}
WI_emerge <- predict_phenology(
  data = woody_island,
  dates = date,
  temperature = temp_c,
  spawn.date = "1990-08-18",
  model = sockeye_emerge_mod   # notice we're using emerge model expression here
  )

# see days to hatch and development period
WI_emerge$days2done
WI_emerge$dev.period
```

@@

```{r}
plot_phenology(WI_hatch)
```

@@

```{r, message=FALSE, echo=FALSE}
cut_ints <- date(
  c(
    "2011-08-31", "2012-08-31",
    "2013-08-31", "2014-08-31",
    "2015-08-31"
  )
)

>>>>

```{r echo = FALSE, out.height = '50%', fig.align = 'center', fig.cap= "Full workflow for hatchR. Data processes are demonstrated with beige circles and pakcage functions with white rectangles."}
knitr::include_graphics("flowchart.png")
```

@@

```{r, fig.cap= "Full output of the plot_check_temp() visual data check function from hatchR. Note that the red and blue lines are customizable by the user to create outlier thresholds for their visual check."}
plot_check_temp(data = year_sim,
                dates = date,
                temperature = temp,
                temp_min = 0, # temp_min and max lines are 
                              # user customizable
                temp_max = 25)
```

@@

@@

@@

Below, we demonstrate how the `fit_model()` function may be used to create custom parameterizations for species beyond the Salmonids in the `model_table` included in the package. We include parameterizations from three warm-water species to demonstrate the `fit_model()` utility for fishes beyond the scope of the original effective value approach. We include parameterizations for commonly cultured sportfishes including Smallmouth Bass (*Micropterus dolomieu*), Channel Catfish (*Ictalurus punctatus*) from @small2001, and Lake Sturgeon (*Acipenser fulvescens*) from @smith2005. We demonstrate parameterization for just Smallmouth Bass here for concision, but the full code set for all species is available at <https://github.com/bmait101/hatchR/blob/master/inst/manuscript/paper.Rmd>.

@@

```{r, echo=FALSE}

@@

```{r, echo=FALSE, warning=FALSE, message=FALSE, fig.cap= "Days to hatch as predicted using custom parameterizations for Channel Catfish (blue), Lake Sturgeon (black, and Smallmouth Bass (green) as generated by the fit_model() function in hathcR. Models are all run using a spawn date starting June 1 for a randomly generated ascending thermal regime with mean temperature 16 °C"}
### predict_phenology

@@

@@

@@

```{r, fig.cap= "Output of the plot_phenology() function from hatchR. The function plots the output of the predict_phenology() function and displays time to develop, as well as the ambient thermal regime (blue), scaled daily effective value (green), and the scaled cumulative sum of effective values (orange)."}
plot_phenology(WI_hatch)
```

@@

```{r, message=FALSE, echo=FALSE, fig.cap= "Densities of predicted days to hatch for Bull Trout for 139 putative spawning locations across Idaho for four years of shared temperature data. Early spawning considers if a parent spawned September 1, Peak-September 15, and Late-September 30."}
cut_ints <- date(
  c(
    "2011-08-31", "2012-08-31",
    "2013-08-31", "2014-08-31",
    "2015-08-31"
  )
)
